Evolution of the AGN UV luminosity function from redshift 7.5

### 2018-03-26 Paper/paper.tex
Update text

\begin{table*}
  % In published version of this table, we could remove file names and add data homogenisation in Notes.
  % Add z ~ 7 qso when you add it to analysis
  % See data.tex for a version of this table that includes file names
  % Should we mark Giallongo, Mortlock, and Banados entries to say that these are sometimes excluded from analysis?
  % Remove notes?
  % Do we really need a sample ID column?
  \caption{Quasar data sets}
  \label{tab:samples}
  \begin{tabular}{crcllrSl}
    \hline
    & ID & $z$ & Survey & Reference & Number & {Area} & Notes \\
    & & & & & of quasars & {(deg$^2$)} & \\
    \hline
    1. & 13 & 0.1--2.2 & SDSS DR7 & \citet{2006AJ....131.2766R} & 48664 & 6248.0 & \\
    2. & 15 & 0.4--2.6 & 2SLAQ SGP & \citet{2009MNRAS.392...19C} & 3663 & 64.2 & \\
    3. & 15  & 0.4--2.6 & --- NGP & \citet{2009MNRAS.392...19C} & 8153 & 127.7 & \\
    4. &  1 & 2.2--3.5 & BOSS DR9 & \citet{2013ApJ...773...14R} & 23301 & 2236.0 & \\
    5. & 13 & 3.7--4.7 & SDSS DR7 & \citet{2006AJ....131.2766R} & 1785 & 6248.0 & Restricted to $z<4.7$ \\
    6. & 17 & 4.7--5.4 & SDSS+WISE & \citet{2016ApJ...829...33Y} & 99 & 14555.0 & Overlaps with 8 for $M_{1450}<-26.73$ \\
    7. &  8 & 4.7--5.1 & SDSS DR7 & \citet{2013ApJ...768..105M} & 148 & 6248.0 & \\
    8. &  8 & 4.7--5.1 & --- Stripe 82 & \citet{2013ApJ...768..105M} & 52 & 235.0 & \\
    9. &  8 & 5.1--5.5 & --- DR7 Extended & \citet{2013ApJ...768..105M} & 28 & 6248.0 & \\
    10. & 8 & 5.1--5.4 & ---  Stripe 82 Extended & \citet{2013ApJ...768..105M} & 10 & 235.0 & \\
    11. & 6 & 3.7--4.8 & NDWFS & \citet{2011ApJ...728L..26G} & 12 & 1.71 & \\
    12. & 6 & 3.8--5.1 & DLS & \citet{2011ApJ...728L..26G} & 12 & 2.05 & \\
    13. & 7 & 4.1--6.3 & CANDELS GOODS-S & \citet{2015AA...578A..83G} & 19 & 0.047 & Rescaled selection probabilities \\
    14. & 18 & 5.8--6.4 & --- Main & \citet{2016ApJ...833..222J} & 24 & 11240.0 & \\
    15. & 18 & 5.9--6.1 & --- Overlap & \citet{2016ApJ...833..222J} & 10 & 4223.0 & \\
    16. & 18 & 5.7--6.1 & --- Stripe 82 & \citet{2016ApJ...833..222J} & 13 & 277.0 & \\
    17. & 10 & 6.0 & CFHQS Deep & \citet{2010AJ....139..906W} & 1 & 4.47 & \\
    18. & 10 & 5.9--6.4 & --- Very Wide & \citet{2010AJ....139..906W} & 16 & 494.0 & \\
    19. & 11 & 6.0--6.2 & Subaru High-$z$ Quasar & \citet{2015ApJ...798...28K} & 2 & 6.5 & \\
    20. & 19 & 6.5--7.4 & UKIDSS & \citet{2011Natur.474..616M} & 1 & 3370.0 & \\
    21. & 19 & 6.5--7.4 & UKIDSS & \citet{2007MNRAS.376L..76V} & 1 & 3370.0 & \\
    22. & 20 & 6.5--7.4 & ALLWISE+UKIDSS+DECaLS & \citet{2018Natur.553..473B} & 1 & 2500.0 & \\
    \hline
  \end{tabular}
\end{table*}

@@

%--------------------

@@

The smooth evolution of the luminosity function parameters apparant in
Figure~\ref{fig:evoln} suggests that it should be possible to describe
the quasar UV luminosity function evolution from $z\sim 7$ to $0$
using far fewer number of parameters.  Such descriptions have been put
forward in the literature for the X-ray and bolometric luminosity
functions \citep[e.g.,][]{2007ApJ...654..731H, 2015MNRAS.451.1892A}.
Such ``global'' descriptions of the luminosity function evolution are
useful as they give a continuous description of the luminosity
function.  This makes them valuable while developing an understanding
of the physics behind the luminosity function. We therefore obtain
estimates of the luminosity function in which the evolution of the
four parameters of the double power law models in
Equation~(\ref{eqn:dpl}) is parameterised to write
\begin{align}
  &\phi_*(z) = f_0(\{c_{0,i}\}, z)\nonumber\\
  &M_*(z) = f_1(\{c_{1,i}\}, z)\nonumber\\
  &\alpha(z) = f_2(\{c_{2,i}\}, z)\nonumber\\
  &\beta(z) = f_3(\{c_{3,i}\}, z),
\label{eqn:global}
\end{align}
where the $c_i$'s are the new model parameters.  The joint posterior
probability distribution of these parameters can be now written as
\begin{equation}
  p(\{c_i\} | \{M_i, z_i\}) \propto p(\{c_i\})p(\{M_i, z_i\} | \{c_i\}),
\end{equation}
where the likelihood 
\begin{equation}
  \mathcal{L}\equiv p(\{M_i, z_i\} | \{c_i\}),
\end{equation}
is now given by $\phi(M,z)$ with suitable normalisation.  Note that
$\phi(M,z)$ is simply 
\begin{multline}
  \phi(M,z) = \phi_*(z) \left[10^{0.4(\alpha(z)+1)(M-M_*(z))}\right. \\ \left.+ 10^{0.4(\beta(z)+1)(M-M_*(z))}\right]^{-1}.
\end{multline}
The ngative logarithm of the likelihood $S\equiv -2\ln\mathcal{L}$ is
straightforward generalisation of Equation~(\ref{eqn:S}), so that
\begin{multline}
  S = -2\sum_{i=1}^N\ln\phi(M_i, z_i)\\+2\int_{M_\mathrm{min}}^{M_\mathrm{max}}dM\int_{z_\mathrm{min}}^{z_\mathrm{max}}dz\, \phi(M,z) f(M, z)\,\frac{dV}{dz},
  \label{eqn:S2}
\end{multline}
where now $N$ is the total number of quasars included in the analysis,
and the integrals in the second term on the right-hand side are over
the complete surveyed range of luminosity and redshift.

We consider models in which the evolution of the four double power law
parameter is modelled independently as in
Equations~(\ref{eqn:global}).  This approach has been termed as
``flexible double power law'' by \citet{2015MNRAS.451.1892A}.  For the
parameters $\phi_*$, $M_*$ and $\alpha$ we assume that the functions
$f_i$s in Equations~(\ref{eqn:global}) are Chebyshev polynomials in
$1+z$.  Thus the functions $f_i$s are written as
\begin{equation}
  f_i=\sum_{j=0}^{n_i}c_{i,j}T_j(1+z),
  \label{eqn:cbs}
\end{equation}
where $c_{i,j}$ are the parameters that appear in
Equations~(\ref{eqn:global}) and $T_j(1+z)$ are Chebyshev polynomials
of the first kind.  We try successively higher orders of Chebyshev
polynomials in order to arrive at a good fit with the data.  We find
that $\phi_*$, $M_*$ and $\alpha$ prefer quadratic, cubic, and linear
evolutions in $1+z$.

The faint-end slope $\beta$ seems to require a break in its evolution
at $z\sim 3$, as discussed in the previous section.  As a result, it
is difficult to model the evolution of $\beta$ using the Chebyshev
expansion of Equation~(\ref{eqn:cbs}).  Instead, we consider a double
power law model for $\beta$ and write
\begin{equation}
  f_3(1+z)=h+\frac{f_0}{10^{a\zeta}+10^{b\zeta}},
\end{equation}
where
\begin{equation}
  \zeta = \log_{10}\left(\frac{1+z}{1+z_0}\right),
\end{equation}
thus resulting in a five-parameter model with parameters $h$, $f_0$,
$z_0$, $a$ and $b$.  The parameters $a$ and $b$ thus determine the low
and high redshift slopes of this evolution, with a break at redshift
$z_0$.  This is similar to the model of \citet{2007ApJ...654..731H},
who also favoured a broken power law model for the evolution of the
faint-end slope of the bolometric luminosity function of quasars.

Our resultant 14-parameter global model for the luminosity function
evolution in shown in Figure~\ref{fig:mosaic_global}, in which the
green curves show a random draw from the posterior probability
distibution function.  For comparison, the symbols show the luminosity
function binned in luminosity and redshift, and the yellow curves show
the posterior distribution of the double power law luminosity
functions in various redshift bins, as in Figure~\ref{fig:mosaic}.  We
find that the global model is in excellent agreement with the binned
models.  The agreement is poor for the BOSS quasars as these are not
described by the global model, as discussed in the previous section.

Figure~\ref{fig:evoln_global} shows the parameter evolution in the
global model, by comparing it with the results from the fits in
individual redshift bins shown in Figure~\ref{fig:evoln}.  The
agreement between the two models is very good, showing an evolution
towards a steep faint-end slope of the luminosity function at high
redshifts, a rapid brightening of the break luminosity.  Note that the
redshift bins with BOSS qsos are not shown in
Figure~\ref{fig:evoln_global}, which explains the difference from
Figure~\ref{fig:evoln}.

The evolution of the comoving number density of quasars is shown in
Figure~\ref{fig:rhoqso} when the luminosity function is integrated
down to different limits.  Similar to Figure~\ref{fig:evoln_global},
symbols show the best fit parameters with 68\% uncertainties from the
double power law fits to the data in redshift bins.  Black solid curve
shows the best fit global model, and the coloured curves show a random
sample drawn from the global posterior.  Bins with BOSS qsos are not
shown.  The number density of qsos down to the limit of spectroscopic
data is about 1 per $10^5$ cMpc$^3$ at its peak.  This density rapidly
increases at low redshifts and then drops gradually at high redshifts.
Figure~\ref{fig:rhoqso} also shows the familiar downsizing feature in
which the number density of faint qsos peaks are lower redshifts than
that of the bright qsos.  While the number density of qsos with
$M_*<-27$ peaks at $z\sim 2.5$, the number density of qsos with
$M_*<-24$ peaks at $z\sim 2$.  At the faintest luminosity at which
spectroscopic data exist, $M_*<-21$, the number density of qsos peaks
at $z\sim 1$.  \gk{Discuss comparison of expected number of quasars at
  redshift 7 with UKIDSS.}

>>>>

\begin{table*}
  % In published version of this table, we could remove file names and add data homogenisation in Notes.
  % Add z ~ 7 qso when you add it to analysis
  % See data.tex for a version of this table that includes file names
  % Should we mark Giallongo, Mortlock, and Banados entries to say that these are sometimes excluded from analysis?
  % Remove notes?
  % Do we really need a sample ID column?
  % Say in caption or text that sample 13 is restricted to $z<4.7$
  % Also that sample 17 overlaps with 8 for $M_{1450}<-26.73$
  % Also that in sample 7 we rescaled selection probabilities 
  \caption{Quasar data sets}
  \label{tab:samples}
  \begin{tabular}{crcllrS}
    \hline
    & ID & $z$ & Survey & Reference & Number & {Area} \\
    & & & & & of quasars & {(deg$^2$)} \\
    \hline
    1. & 13 & 0.1--2.2 & SDSS DR7 & \citet{2006AJ....131.2766R} & 48664 & 6248.0 \\
    2. & 15 & 0.4--2.6 & 2SLAQ SGP & \citet{2009MNRAS.392...19C} & 3663 & 64.2 \\
    3. & 15  & 0.4--2.6 & --- NGP & \citet{2009MNRAS.392...19C} & 8153 & 127.7 \\
    4. &  1 & 2.2--3.5 & BOSS DR9 & \citet{2013ApJ...773...14R} & 23301 & 2236.0 \\
    5. & 13 & 3.7--4.7 & SDSS DR7 & \citet{2006AJ....131.2766R} & 1785 & 6248.0 \\
    6. & 17 & 4.7--5.4 & SDSS+WISE & \citet{2016ApJ...829...33Y} & 99 & 14555.0 \\
    7. &  8 & 4.7--5.1 & SDSS DR7 & \citet{2013ApJ...768..105M} & 148 & 6248.0 \\
    8. &  8 & 4.7--5.1 & --- Stripe 82 & \citet{2013ApJ...768..105M} & 52 & 235.0 \\
    9. &  8 & 5.1--5.5 & --- DR7 Extended & \citet{2013ApJ...768..105M} & 28 & 6248.0 \\
    10. & 8 & 5.1--5.4 & ---  Stripe 82 Extended & \citet{2013ApJ...768..105M} & 10 & 235.0 \\
    11. & 6 & 3.7--4.8 & NDWFS & \citet{2011ApJ...728L..26G} & 12 & 1.71 \\
    12. & 6 & 3.8--5.1 & DLS & \citet{2011ApJ...728L..26G} & 12 & 2.05 \\
    13. & 7 & 4.1--6.3 & CANDELS GOODS-S & \citet{2015AA...578A..83G} & 19 & 0.047 \\
    14. & 18 & 5.8--6.4 & --- Main & \citet{2016ApJ...833..222J} & 24 & 11240.0 \\
    15. & 18 & 5.9--6.1 & --- Overlap & \citet{2016ApJ...833..222J} & 10 & 4223.0 \\
    16. & 18 & 5.7--6.1 & --- Stripe 82 & \citet{2016ApJ...833..222J} & 13 & 277.0 \\
    17. & 10 & 6.0 & CFHQS Deep & \citet{2010AJ....139..906W} & 1 & 4.47 \\
    18. & 10 & 5.9--6.4 & --- Very Wide & \citet{2010AJ....139..906W} & 16 & 494.0 \\
    19. & 11 & 6.0--6.2 & Subaru High-$z$ Quasar & \citet{2015ApJ...798...28K} & 2 & 6.5 \\
    20. & 19 & 6.5--7.4 & UKIDSS & \citet{2011Natur.474..616M} & 1 & 3370.0 \\
    21. & 19 & 6.5--7.4 & UKIDSS & \citet{2007MNRAS.376L..76V} & 1 & 3370.0 \\
    22. & 20 & 6.5--7.4 & ALLWISE+UKIDSS+DECaLS & \citet{2018Natur.553..473B} & 1 & 2500.0 \\
    \hline
  \end{tabular}
\end{table*}

@@

\begin{figure*}
  \begin{center}
    \begin{tabular}{cc}
      \includegraphics[width=0.47\textwidth]{emissivity_18.pdf} & 
      \includegraphics[width=0.47\textwidth]{emissivity_21.pdf} \\
      % gammapi.py -- draw_emissivity()
    \end{tabular}
  \end{center}
  \caption{The 912~\AA\ emissivity of AGN assuming 100\% escape
    fraction down to a limiting magnitude.  Brown points show the
    emissivity when the best-fit double power law luminosity functions
    at respective redshifts are integrated down to $M_{1450}=-18$.
    Black points show the integration limit when the limit is assumed
    to be $M_{1450}=-21$.  For comparison, various other published
    models of the 1~Ry emissivity of AGN are also shown.  The blue
    curve in these panels shows the AGN 912~\AA\ emissivity model of
    \citet{2012ApJ...746..125H}, which is very similar to the model of
    \citet{2007ApJ...654..731H}.  Red circles with error bars show the
    determination from the fits of \citet{2015AA...578A..83G}.  The
    maroon curve is the published fit to the 1~Ry emissivity evolution
    in the fits of \citet{2017MNRAS.466.1160M}.  The green curve is
    the model of \citet{2015ApJ...813L...8M}.  The grey and brown
    curves show the best-fit form of Equation XX. \gk{Mention MH15 and
      HM12 Lstar limits.  Mention that Akiyama open and filled circles
      are very close.}}
  \label{fig:e912_2}
\end{figure*}

@@

The smooth evolution of the luminosity function parameters apparant in
Figure~\ref{fig:evoln} suggests that it should be possible to describe
the quasar UV luminosity function evolution out to $z\sim 7$ using far
fewer number of parameters.  Such descriptions have been developed in
the literature for the X-ray \citep[e.g.,][]{2015MNRAS.451.1892A} and
bolometric luminosity functions \citep[e.g.,][]{2007ApJ...654..731H}.
Such ``global'' descriptions of the luminosity function evolution are
useful as they give a continuous description of the luminosity
function.  This allows one to reduce the bias introduced by binning
the data in arbitrary redshift bins.  By potentially allowing for
extrapolations beyond the redshift range spanned by the data, such
models are valuable for understanding of the physics behind the
luminosity function.  Ideally, one would want to use physically
meaningful parameters that govern the formation and evolution of the
AGN population.  Unfortunately, such physical parameterisation is yet
to be developed.  We therefore set up an empirical parameterisation to
describe the evolution of the four parameters of the double power law
model in Equation~(\ref{eqn:dpl}) as 
\begin{align}
  &\phi_*(z) = f_0(\{c_{0,i}\}, z)\nonumber\\
  &M_*(z) = f_1(\{c_{1,i}\}, z)\nonumber\\
  &\alpha(z) = f_2(\{c_{2,i}\}, z)\nonumber\\
  &\beta(z) = f_3(\{c_{3,i}\}, z),
\label{eqn:global}
\end{align}
where the $c_{j,i}$'s are the new model parameters, and the $f_j$'s
are functions that vary smoothly with redshift $z$.  The joint
posterior probability distribution of these parameters can be now
written as
\begin{equation}
  p(\{c_i\} | \{M_i, z_i\}) \propto p(\{c_i\})p(\{M_i, z_i\} | \{c_i\}),
\end{equation}
where the likelihood 
\begin{equation}
  \mathcal{L}\equiv p(\{M_i, z_i\} | \{c_i\}),
\end{equation}
is now given by $\phi(M,z)$ with suitable normalisation.  Note that
$\phi(M,z)$ is given by Equation~(\ref{eqn:dpl}), but now the four
parameters in that equation are redshift-dependent, so that
\begin{multline}
  \phi(M,z) = \phi_*(z) \left[10^{0.4(\alpha(z)+1)(M-M_*(z))}\right. \\ \left.+ 10^{0.4(\beta(z)+1)(M-M_*(z))}\right]^{-1}.
\end{multline}
The negative logarithm of the likelihood $S\equiv -2\ln\mathcal{L}$ is
straightforward generalisation of Equation~(\ref{eqn:S}), given by 
\begin{multline}
  S = -2\sum_{i=1}^N\ln\phi(M_i, z_i)\\+2\int_{M_\mathrm{min}}^{M_\mathrm{max}}dM\int_{z_\mathrm{min}}^{z_\mathrm{max}}dz\, \phi(M,z) f(M, z)\,\frac{dV}{dz},
  \label{eqn:S2}
\end{multline}
where now $N$ is the total number of quasars included in the analysis,
and the integrals in the second term on the right-hand side are over
the complete surveyed range of luminosity and redshift, instead of
redshift and luminosity bins.  We consider models in which the
evolution of the four double power law parameter is modelled
independently as in Equations~(\ref{eqn:global}).  This approach has
been termed as ``flexible double power law'' by
\citet{2015MNRAS.451.1892A}.  We present three such models in this
paper.  These are shown in Figure~\ref{fig:evoln_global}.  The three
models differ in the way they describe the evolution of the faint-end
slope $\beta$ and in the selection of the AGN data.  The models are
defined as follows.

\begin{itemize}

\item In Model 1, the parameters $\phi_*$, $M_*$ and $\alpha$ we
  assume that the functions $f_i$'s in Equations~(\ref{eqn:global})
  are Chebyshev polynomials in $(1+z)$.  Thus the functions $f_i$s are
  written as
  \begin{equation}
    f_i=\sum_{j=0}^{n_i}c_{i,j}T_j(1+z),
    \label{eqn:cbs}
  \end{equation}
  where $c_{i,j}$ are the parameters that appear in
  Equations~(\ref{eqn:global}) and $T_j(1+z)$ are Chebyshev
  polynomials of the first kind.  We try successively higher orders of
  Chebyshev polynomials in order to arrive at a good fit with the
  data.  As we discuss below, we find that $\phi_*$, $M_*$ and
  $\alpha$ prefer quadratic, cubic, and linear evolutions in $(1+z)$.
  The faint-end slope $\beta$ seems to require a break in its
  evolution at $z\sim 3$, as discussed in the previous section.  As a
  result, it is difficult to model the evolution of $\beta$ using the
  Chebyshev expansion of Equation~(\ref{eqn:cbs}).  Instead, we
  consider a double power law model for $\beta$ and write \gk{Is this
    really only four parameters?}
  \begin{equation}
    f_3(1+z)=c_{3,0}+\frac{c_{3,1}}{10^{c_{3,3}\zeta}+10^{c_{3,4}\zeta}},
  \end{equation}
  where
  \begin{equation}
    \zeta = \log_{10}\left(\frac{1+z}{1+c_{3,2}}\right),
  \end{equation}
  thus resulting in a five-parameter model with parameters $c_{3,i}$.
  The parameters $c_{3,3}$ and $c_{3,4}$ thus determine the low and
  high redshift slopes of this evolution, with a break at redshift
  $c_{3,2}$.  This is similar to the model of
  \citet{2007ApJ...654..731H}, who also favoured a broken power law
  model for the evolution of the faint-end slope of the bolometric
  luminosity function of quasars.  Model 1 thus has 14 parameters.
  Excluding those deemed to be dominated by systematic errors, as
  discussed in the previous section, all of the remaining AGN from
  Table~\ref{tab:samples} are included while fitting this model.

\item Model 2 parameterises the luminosity function evolution in the
  same way as Model 1, so that the faint-end slope evolution is
  described by a double power law while the evolution of the other
  parameters $\phi_*$, $M_*$ and $\alpha$ is described by,
  respectively, quadratic, cubic, and linear polynomials in $(1+z)$.
  The total number of parameter is 14.  However, while fitting this
  model, we drop samples 7, 19, and 20 from the analysis.  The
  selection function for these samples are not yet well-characterised.
  Removing them allows us to understand the effect this has on the
  favoured evolution model.

\item In Model 3, we again exclude samples 7, 19, and 20 from the
  analysis.  We also continue to describe the evolution of $\phi_*$,
  $M_*$ and $\alpha$ by quadratic, cubic, and linear polynomials in
  $(1+z)$, respectively.  But in this model, the evolution of the
  faint-end slope $\beta$ is also assumed to be linear in $(1+z)$.
  This model thus has just 11 parameters.
\end{itemize}

Figure~\ref{fig:mosaic_global} shows the three models in comparison
with the binned fits from the previous section.  The shaded regions
show the one-sigma (68.26\%) uncertainty.  The symbols show the
luminosity function binned in luminosity and redshift, and the yellow
shaded regions show the posterior distribution of the double power law
luminosity functions in various redshift bins, as in
Figure~\ref{fig:mosaic}.  Bins containing data with large systematic
error are excluded from Figure~\ref{fig:mosaic_global}.  All three
global models are in excellent agreement with the binned models,
although Model~1 performs better.

Figure~\ref{fig:evoln_global} shows the parameter evolution in the
global models, by comparing it with the results from the fits in
individual redshift bins shown in Figure~\ref{fig:evoln}.  All three
models capture the steepening of the faint end of the luminosity
function towards higher redshifts.  The best-fit form of Model~1 is
shown by the black curves in Figure~\ref{fig:evoln_global}.  The
accompanying grey shaded area depicts the one-sigma (68.26\%)
uncertainty.  The model is in excellent agreement with the results of
the fits in redshift bins discussed in the previous section.  The
deviation of the BOSS quasars at $z=2$--$4$ from the smooth evolution
is again strikingly apparent.  So is the deviation of the SDSS and
2SLAQ quasars at $z<0.6$.  This is an indirect justification of the
data selection discussion previously in Section~\ref{sec:bins}.
Unfortunately, this model suffers with a remarkably sharp break in the
evolution of the faint-end slope $\beta$ at about $z\sim 3.5$.  As
seen in Figure~\ref{fig:evoln_global}, the data seem to require this
break, although it seems unlikely that such a sharp break at this
redshift would be physical.  Model 1 thus serves to emphasize the
necessity of better quality data at these redshifts.  Models 2 and 3
are shown in Figure~\ref{fig:evoln_global} by the green and orange
curves, respectively.  The corresponding best-fit parameter values are
tabulated in Table~\ref{tab:global}.

The evolution of the comoving number density of quasars is shown in
Figure~\ref{fig:rhoqso} when the luminosity function is integrated
down to different limits.  Similar to Figure~\ref{fig:evoln_global},
symbols show the best-fit values with 68\% uncertainties from the
double power law fits to the data in redshift bins.  Black solid curve
shows the best fit global model, and the coloured curves show a random
sample drawn from the global posterior.  This number density evolution
again highlights the systematic error in data at $z\sim 3$.  The
number density of AGN down to the limit of spectroscopic data
($M_{1450}<-21$) is about $10^{-5}$ cMpc$^{-3}$ at its peak.  This
density rapidly increases at low redshifts and then drops gradually at
high redshifts.  Figure~\ref{fig:rhoqso} also shows the familiar
downsizing feature in which the number density of faint qsos peaks are
lower redshifts than that of the bright qsos.  While the number
density of qsos with $M_*<-27$ peaks at $z\sim 2.5$, the number
density of qsos with $M_*<-24$ peaks at $z\sim 2$.  At the faintest
luminosity at which spectroscopic data exist, $M_*<-21$, the number
density of qsos peaks at $z\sim 1$.  However, the difference between
our three models is dramatically evidencet in Figure~\ref{fig:rhoqso}.
Model~1 prefers a decrease in the number density of faint quasars at
$z\sim 3$ followed by an increase at higher redshift.  This is caused
by the rapid steepening of the faint-end slope in this model at this
redshift.  Figure~\ref{fig:rhoqso} reveals another property of these
models: when extrapolated, the AGN number density diverges in all
three models are high redshifts.  This is result of the steep
faint-end slope at high redshifts combined with the rapid brightening
of the break luminosity at $z\sim 3$.  While no data exist at redshift
$z>7.5$, this divergent behaviour is shared by previous models in the
literature.  Figure~\ref{fig:rhoqso} also shows that although Models 2
and 3 exhibit regular behaviour in the evolution of the AGN number
density at $z\sim 3$, they do not fit the $z\sim 4$--$5$ data as well
as Model 1.  \gk{Discuss comparison of expected number of quasars at
  redshift 7 with UKIDSS.  Also discuss our new fit?}
